---
title: "Chapter8 - User feedback"
output: html_notebook
---


```{r setup}
knitr::opts_chunk$set(eval = FALSE)
```


```{r}
library(shiny)
```

# 8.1 Validation

## 8.1.1 Validating input

you should add `useShinyFeedback()` to the ui and then use `feedback()`, `feedbackWarning()`, `feedbackDanger()`, or `feedbackSuccess()` in `server`.

these 4 functions have 3 main arguments and 2 optional arguments for customization
1- `inputid`: the id of the input where the feedback should be placed.
2- `show`: a logical expression, determining whether or not to show the feedback.
3- `text`: the text for the user.
* `icon`
* `color`

```{r}
ui <- fluidPage(
  shinyFeedback::useShinyFeedback(),
  numericInput("n", "n", value = 10),
  textOutput("half")
)

server <- function(input, output, session) {
  observeEvent(input$n,
    shinyFeedback::feedbackWarning(
      "n", 
      input$n %% 2 != 0,
      "Please select an even number"
    )  
  )
  output$half <- renderText(input$n / 2)
}
shinyApp(ui, server)
```


To stop inputs from triggering reactive changes, use `req()` that takes a logical value.

```{r}
server <- function(input, output, session) {
  half <- reactive({
    even <- input$n %% 2 == 0
    shinyFeedback::feedbackWarning("n", !even, "Please select an even number")
    req(even)
    input$n / 2    
  })
  
  output$half <- renderText(half())
}
shinyApp(ui, server)
```


## 8.1.2 Pausing execution with req()

you can stop an action by setting the default `value` to "" (either In `textInput()` by using `value = ""` or by  `selectInput()`, by provide an empty choice, `""`).

```{r}
ui <- fluidPage(
  selectInput("language", "Language", choices = c("", "English", "Maori")),
  textInput("name", "Name"),
  textOutput("greeting")
)

server <- function(input, output, session) {
  greetings <- c(
    English = "Hello", 
    Maori = "Ki ora"
  )
  output$greeting <- renderText({
    paste0(greetings[[input$language]], " ", input$name, "!")
  })
}
shinyApp(ui, server)
```

but in this way you got an error, so you can solve this problem by using `req()`:

```{r}
server <- function(input, output, session) {
  greetings <- c(
    English = "Hello", 
    Maori = "Ki ora"
  )
  output$greeting <- renderText({
    req(input$language, input$name)
    paste0(greetings[[input$language]], " ", input$name, "!")
  })
}
shinyApp(ui, server)
```


## 8.1.3 req() and validation


using `cancelOutput = TRUE` leaves them displaying the last good value.
```{r}
ui <- fluidPage(
  shinyFeedback::useShinyFeedback(),
  textInput("dataset", "Dataset name"), 
  tableOutput("data")
)

server <- function(input, output, session) {
  data <- reactive({
    req(input$dataset)
    
    exists <- exists(input$dataset, "package:datasets")
    shinyFeedback::feedbackDanger("dataset", !exists, "Unknown dataset")
    req(exists, cancelOutput = TRUE)

    get(input$dataset, "package:datasets")
  })
  
  output$data <- renderTable({
    head(data())
  })
}
shinyApp(ui, server)
```

## 8.1.4 Validate output
 sometimes the invalid state is a result of a combination of inputs so you want to show the error in the output.
 so you must use `shiny::validate()`:
 
```{r}
ui <- fluidPage(
  numericInput("x", "x", value = 0),
  selectInput("trans", "transformation", 
    choices = c("square", "log", "square-root")
  ),
  textOutput("out")
)

server <- function(input, output, server) {
  output$out <- renderText({
    if (input$x < 0 && input$trans %in% c("log", "square-root")) {
      validate("x can not be negative for this transformation")
    }
    
    switch(input$trans,
      square = input$x ^ 2,
      "square-root" = sqrt(input$x),
      log = log(input$x)
    )
  })
}
shinyApp(ui, server)
```


# 8.2 Notifications

## 8.2.1 Transient notification

```{r}
ui <- fluidPage(
  actionButton("goodnight", "Good night")
)
server <- function(input, output, session) {
  observeEvent(input$goodnight, {
    showNotification("So long")
    Sys.sleep(1)
    showNotification("Farewell")
    Sys.sleep(1)
    showNotification("Auf Wiedersehen")
    Sys.sleep(1)
    showNotification("Adieu")
  })
}
shinyApp(ui, server)
```

By default, the message will disappear after 5 seconds.
use `duration` argument to override this default.

```{r}
server <- function(input, output, session) {
  observeEvent(input$goodnight, {
    showNotification("So long", duration = 1)
    Sys.sleep(1)
    showNotification("Farewell", type = "message")
    Sys.sleep(1)
    showNotification("Auf Wiedersehen", type = "warning")
    Sys.sleep(1)
    showNotification("Adieu", type = "error")
  })
}
shinyApp(ui, server)
```

## 8.2.2 Removing on completion

```{r}
server <- function(input, output, session) {
  data <- reactive({
    id <- showNotification("Reading data...", duration = NULL, closeButton = FALSE)
    on.exit(removeNotification(id), add = TRUE)
    
    read.csv(input$file$datapath)
  })
}
```


## 8.2.3 Progressive updates

```{r}
ui <- fluidPage(
  tableOutput("data")
)

server <- function(input, output, session) {
  notify <- function(msg, id = NULL) {
    showNotification(msg, id = id, duration = NULL, closeButton = FALSE)
  }

  data <- reactive({ 
    id <- notify("Reading data...")
    on.exit(removeNotification(id), add = TRUE)
    Sys.sleep(1)
      
    notify("Reticulating splines...", id = id)
    Sys.sleep(1)
    
    notify("Herding llamas...", id = id)
    Sys.sleep(1)

    notify("Orthogonalizing matrices...", id = id)
    Sys.sleep(1)
        
    mtcars
  })
  
  output$data <- renderTable(head(data()))
}
shinyApp(ui, server)
```


# 8.3 Progress bars
## 8.3.1 Shiny

```{r}
# Create a progress bar object with `Progress$new(max = number_of_steps)`.
progress <- Progress$new(max = 5)

# Display the progress bar by calling the `$set()` method, 
# providing a title for the progress bar in the `message` argument.
progress$set(message = "Starting process")

# Call `$inc()` repeatedly, once for each step.
for (i in 1:5) {
  progress$inc(1)
}

# When done, call `$close()` to terminate the progress bar.
progress$close()
```

```{r}
ui <- fluidPage(
  numericInput("steps", "How many steps?", 10),
  actionButton("go", "go"),
  textOutput("result")
)

server <- function(input, output, session) {
  data <- reactive({
    req(input$go)
    
    progress <- Progress$new(max = input$steps)
    on.exit(progress$close())
    
    progress$set(message = "Computing random number")
    for (i in seq_len(input$steps)) {
      Sys.sleep(0.5)
      progress$inc(1)
    }
    runif(1)
  })
  
  output$result <- renderText(round(data(), 2))
}
shinyApp(ui, server)
```


## 8.3.2 Waiter

```{r}
ui <- fluidPage(
  waiter::use_waitress(),
  numericInput("steps", "How many steps?", 10),
  actionButton("go", "go"),
  textOutput("result")
)
server <- function(input, output, session) {
  data <- reactive({
    req(input$go)
    waitress <- waiter::Waitress$new(max = input$steps)
    on.exit(waitress$close())
    
    for (i in seq_len(input$steps)) {
      Sys.sleep(0.5)
      waitress$inc(1)
    }
    
    runif(1)
  })
  
  output$result <- renderText(round(data(), 2))
}
shinyApp(ui, server)
```



You can override the default `theme` to use one of:

`overlay`: an opaque progress bar that hides the whole page
`overlay-opacity`: a translucent progress bar that covers the whole page
`overlay-percent`: an opaque progress bar that also displays a numeric percentage.

Instead of showing a progress bar for the entire page, you can overlay it on an existing input or output by setting the `selector` parameter, e.g.:
```{r}
ui <- fluidPage(
  waiter::use_waitress(),
  numericInput("steps", "How many steps?", 10),
  actionButton("go", "go"),
  textOutput("result")
)
server <- function(input, output, session) {
  data <- reactive({
    req(input$go)
    waitress <- waiter::Waitress$new(selector = "#steps", theme = "overlay")
    on.exit(waitress$close())
    
    for (i in seq_len(input$steps)) {
      Sys.sleep(0.5)
      waitress$inc(1)
    }
    
    runif(1)
  })
  
  output$result <- renderText(round(data(), 2))
}
shinyApp(ui, server)

```

## 8.3.3 Spinners
if you don’t know exactly how long an operation will take, use `waiter` instead of `waitress`
```{r}
ui <- fluidPage(
  waiter::use_waiter(),
  actionButton("go", "go"),
  textOutput("result")
)

server <- function(input, output, session) {
  data <- reactive({
    req(input$go)
    waiter <- waiter::Waiter$new()
    waiter$show()
    on.exit(waiter$hide())
    
    Sys.sleep(sample(5, 1))
    runif(1)
  })
  output$result <- renderText(round(data(), 2))
}
shinyApp(ui, server)
```


```{r}
ui <- fluidPage(
  waiter::use_waiter(),
  actionButton("go", "go"),
  plotOutput("plot"),
)

server <- function(input, output, session) {
  data <- reactive({
    req(input$go)
    waiter::Waiter$new(id = "plot")$show()
    
    Sys.sleep(3)
    data.frame(x = runif(50), y = runif(50))
  })
  
  output$plot <- renderPlot(plot(data()), res = 96)
}
shinyApp(ui, server)
```

An even simpler alternative is to use the `shinycssloaders` package. you just use `shinycssloaders::withSpinner()` to wrap outputs that you want to automatically get a spinner when they have been invalidated.

```{r}
library(shinycssloaders)

ui <- fluidPage(
  actionButton("go", "go"),
  withSpinner(plotOutput("plot")),
)
server <- function(input, output, session) {
  data <- reactive({
    req(input$go)
    Sys.sleep(3)
    data.frame(x = runif(50), y = runif(50))
  })
  
  output$plot <- renderPlot(plot(data()), res = 96)
}
shinyApp(ui, server)
```

# 8.4 Confirming and undoing

## 8.4.1 Explicit confirmation

```{r}
modal_confirm <- modalDialog(
  "Are you sure you want to continue?",
  title = "Deleting files",
  footer = tagList(
    actionButton("cancel", "Cancel"),
    actionButton("ok", "Delete", class = "btn btn-danger")
  )
)
modal_confirm
```


example:
```{r}
ui <- fluidPage(
  actionButton("delete", "Delete all files?")
)
server <- function(input, output, session) {
  observeEvent(input$delete, {
    showModal(modal_confirm)
  })
  
  observeEvent(input$ok, {
    showNotification("Files deleted")
    removeModal()
  })
  observeEvent(input$cancel, 
    removeModal()
  )
}
shinyApp(ui, server)
```


## 8.4.2 Undoing an action

```{r}
ui <- fluidPage(
  textAreaInput("message", 
    label = NULL, 
    placeholder = "What's happening?",
    rows = 3
  ),
  actionButton("tweet", "Tweet")
)

runLater <- function(action, seconds = 3) {
  observeEvent(
    invalidateLater(seconds * 1000), action, 
    ignoreInit = TRUE, 
    once = TRUE, 
    ignoreNULL = FALSE,
    autoDestroy = FALSE
  )
}

server <- function(input, output, session) {
  waiting <- NULL
  last_message <- NULL
  
  observeEvent(input$tweet, {
    notification <- glue::glue("Tweeted '{input$message}'")
    last_message <<- input$message
    updateTextAreaInput(session, "message", value = "")

    showNotification(
      notification,
      action = actionButton("undo", "Undo?"),
      duration = NULL,
      closeButton = FALSE,
      id = "tweeted",
      type = "warning"
    )

    waiting <<- runLater({
      cat("Actually sending tweet...\n")
      removeNotification("tweeted")
    })
  })
  
  observeEvent(input$undo, {
    waiting$destroy()
    showNotification("Tweet retracted", id = "tweeted")
    updateTextAreaInput(session, "message", value = last_message)
  })
}
shinyApp(ui, server)
```


