---
title: "Chapter2-Your first Shiny app"
output: html_notebook
---
```{r setup}
knitr::opts_chunk$set(eval = FALSE)
```


```{r}
library(shiny)
```


write "shinyapp" and then press shift + tab to use its snippets.

# 2.2 Create app directory and file

```{r}
file.edit("app.R")
```

```{r}
# two way to run an shiny app
# 1
source("app.R")
# 2
runApp()
```

```{r}
ui <- fluidPage(
  "Hello, world!"
)
server <- function(input, output, session) {
}
shinyApp(ui, server)
```

# 2.4 Adding UI controls

```{r}

ui <- fluidPage(
  selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)
server <- function(input, output, session) {
}
shinyApp(ui, server)
```   


# 2.5 Adding behavior

```{r}
ui <- fluidPage(
  selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)

server <- function(input, output, session) {
  output$summary <- renderPrint({
    dataset <- get(input$dataset, "package:datasets")
    summary(dataset)
  })
  
  output$table <- renderTable({
    dataset <- get(input$dataset, "package:datasets")
    dataset
  })
}
shinyApp(ui, server)
```

we must give shiny the recipe for making the outputs and the pattern is like this:

```{r}
output$ID <- renderTYPE({
  # Expression that generates whatever kind of output
  # renderTYPE expects
})
```

the render function `render*` and the output control of input `*Output` must be the same type.
in this case we use `renderTable` to display the selected input and the render type matches with `tableOutput`.



# 2.6 Reducing duplication with reactive expressions

we can stop the duplication in code (`dataset <- get(input$dataset, "package:datasets")`) by a technique called "reactive expression" using `reactive()`.
this technique retrieve the input once and caches the result until it needs to be updated.


```{r}
ui <- fluidPage(
  selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)

server <- function(input, output, session) {
  dataset <- reactive({
    get(input$dataset, "package:datasets")
  })

  output$summary <- renderPrint({
    summary(dataset())
  })
  
  output$table <- renderTable({
    dataset()
  })
}

shinyApp(ui, server)
```


# 2.8 Exercises

question 1:
```{r}
library(shiny)

ui <- fluidPage(textInput("name", "What's your name?"), textOutput("greeting"))

server <- function(input, output, session) {
  output$greeting <- renderText({
  paste0("Hello ", input$name)
})
}

shinyApp(ui, server)

```

question 2:

```{r}

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  "then x times 5 is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * 5
  })
}
shinyApp(ui, server)
```



question 3:

```{r}

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", label = "and y is", min = 1, max = 50, value = 30),
  "then x times y is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * input$y
  })
}
shinyApp(ui, server)
```


question 4:

```{r}
ui <- fluidPage(
  sliderInput("x", "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", "and y is", min = 1, max = 50, value = 5),
  "then, (x * y) is", textOutput("product"),
  "and, (x * y) + 5 is", textOutput("product_plus5"),
  "and (x * y) + 10 is", textOutput("product_plus10")
)

server <- function(input, output, session) {
  product <- reactive({ 
    product <- input$x * input$y
  })
  output$product <- reactive({ 
    product()
  })
  output$product_plus5 <- renderText({ 
    product() + 5
  })
  output$product_plus10 <- renderText({ 
    product() + 10
  })
}
shinyApp(ui, server)
```


question 5:
```{r}
library(ggplot2)
datasets <- data(package = "ggplot2")$results[c(2, 4, 10), "Item"]

ui <- fluidPage(
  selectInput("dataset", "Dataset", choices = datasets),
  verbatimTextOutput("summary"),
  plotOutput("plot")
)

server <- function(input, output, session) {
  dataset <- reactive({
    get(input$dataset, "package:ggplot2")
  })
  output$summry <- renderPrint({
    summary(dataset())
  })
  output$plot <- renderPlot({
    plot(dataset())
  }, res = 96)
}
shinyApp(ui, server)
```