---
title: "Chapter16 - Coordinate systems"
output: html_notebook
---

# 16.1 Linear coordinate systems

There are three linear coordinate systems: `coord_cartesian()`, `coord_flip()`, `coord_fixed()`.


## 16.1.1 Zooming into a plot with coord_cartesian()

```{r}
base <- ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth()

# Full dataset
base

# Scaling to 4--6 throws away data outside that range
base + scale_x_continuous(limits = c(4, 6))

base + coord_cartesian(xlim = c(4, 6))
```

## 16.1.2 Flipping the axes with coord_flip()

```{r}
ggplot(mpg, aes(displ, cty)) + 
  geom_point() + 
  geom_smooth()

# Exchanging cty and displ rotates the plot 90 degrees, but the smooth 
# is fit to the rotated data.
ggplot(mpg, aes(cty, displ)) + 
  geom_point() + 
  geom_smooth()

# coord_flip() fits the smooth to the original data, and then rotates 
# the output
ggplot(mpg, aes(displ, cty)) + 
  geom_point() + 
  geom_smooth() + 
  coord_flip()
```

## 16.1.3 Equal scales with coord_fixed()
`coord_fixed()` fixes the ratio of length on the x and y axes.

# 16.2 Non-linear coordinate systems

```{r}
rect <- data.frame(x = 50, y = 50)
line <- data.frame(x = c(1, 200), y = c(100, 1))
base <- ggplot(mapping = aes(x, y)) + 
  geom_tile(data = rect, aes(width = 50, height = 50)) + 
  geom_line(data = line) + 
  xlab(NULL) + ylab(NULL)
base
base + coord_polar("x")
base + coord_polar("y")


base + coord_flip()
base + coord_trans(y = "log10")
base + coord_fixed()
```



```{r}
df <- data.frame(r = c(0, 1), theta = c(0, 3 / 2 * pi))
ggplot(df, aes(r, theta)) + 
  geom_line() + 
  geom_point(size = 2, colour = "red")
```

```{r}
interp <- function(rng, n) {
  seq(rng[1], rng[2], length = n)
}
munched <- data.frame(
  r = interp(df$r, 15),
  theta = interp(df$theta, 15)
)

ggplot(munched, aes(r, theta)) + 
  geom_line() + 
  geom_point(size = 2, colour = "red")
```

```{r}
transformed <- transform(munched,
  x = r * sin(theta),
  y = r * cos(theta)
)

ggplot(transformed, aes(x, y)) + 
  geom_path() + 
  geom_point(size = 2, colour = "red") + 
  coord_fixed()
```

## 16.2.1 Transformations with coord_trans()


```{r}
# Linear model on original scale is poor fit
base <- ggplot(diamonds, aes(carat, price)) + 
  stat_bin2d() + 
  geom_smooth(method = "lm") + 
  xlab(NULL) + 
  ylab(NULL) + 
  theme(legend.position = "none")
base

# Better fit on log scale, but harder to interpret
base +
  scale_x_log10() + 
  scale_y_log10()

# Fit on log scale, then backtransform to original.
# Highlights lack of expensive diamonds with large carats
pow10 <- scales::exp_trans(10)
base +
  scale_x_log10() + 
  scale_y_log10() + 
  coord_trans(x = pow10, y = pow10)
```

## 16.2.2 Polar coordinates with coord_polar()

```{r}
base <- ggplot(mtcars, aes(factor(1), fill = factor(cyl))) +
  geom_bar(width = 1) + 
  theme(legend.position = "none") + 
  scale_x_discrete(NULL, expand = c(0, 0)) +
  scale_y_continuous(NULL, expand = c(0, 0))

# Stacked barchart
base

# Pie chart
base + coord_polar(theta = "y")

# The bullseye chart
base + coord_polar()
```

## 16.2.3 Map projections with coord_map()

```{r}
# Prepare a map of NZ
nzmap <- ggplot(map_data("nz"), aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  xlab(NULL) + ylab(NULL)

# Plot it in cartesian coordinates
nzmap
# With the aspect ratio approximation
nzmap + coord_quickmap()
```



```{r}
world <- map_data("world")
worldmap <- ggplot(world, aes(long, lat, group = group)) +
  geom_path() +
  scale_y_continuous(NULL, breaks = (-2:3) * 30, labels = NULL) +
  scale_x_continuous(NULL, breaks = (-4:4) * 45, labels = NULL)

worldmap + coord_map()
# Some crazier projections
worldmap + coord_map("ortho")
worldmap + coord_map("stereographic")
```

