---
title: "Diamond sizes"
output: html_document
date: '2016-08-25'
---


Checking Name (and include option)
----------------------------------------------
```{r setup, include = FALSE}
library(tidyverse)
print("i will run before any other codes(becuase of my name)")
```
use the primary object name for chank name


Options
---------------------------------------------


### eval option

```{r eval = FALSE}
test <- "eval works like commenting"
```

### echo option
```{r echo = FALSE}
summary(mtcars)
```



### message and warning option
```{r message = FALSE}
message("hello?")
```


```{r warning = FALSE}
warning("bye!")
```


### results option
```{r results = 'hide'}
summary(mtcars)
```



### fig.show option
```{r fig.show = 'hide'}
plot(pressure)
```



### error = TRUE option
this option good for educational purposes, because rendering won't pause even after an error


Tables
----------------------------------------
this is a normal output:
```{r}
mtcars[1:5, ]
```
but if we want some formatting we need to use **knitr::kable** :
```{r}
knitr::kable(
  mtcars[1:5, ], 
  caption = "A knitr kable."
)

df <- tibble::tibble(c(T,T,F,T,F,F,F))
knitr::kable(df)
```


For even deeper customisation, consider the **xtable, stargazer, pander, tables, and ascii** packages.



Caching
--------------------------
```{r raw_data}
rawdata <- readr::read_csv("a_very_large_file.csv")
```

```{r processed_data, cache = TRUE}
processed_data <- rawdata %>% 
  filter(!is.na(import_var)) %>% 
  mutate(new_variable = complicated_transformation(x, y, z))
```


Caching the processed_data chunk means that it will get re-run if the dplyr pipeline is changed, but it won’t get rerun if the read_csv() call changes. You can avoid that problem with the dependson chunk option:

```{r processed_data, cache = TRUE, dependson = "raw_data"}
processed_data <- rawdata %>% 
  filter(!is.na(import_var)) %>% 
  mutate(new_variable = complicated_transformation(x, y, z))
```

dependson should contain a character vector of every chunk that the cached chunk depends on. Knitr will update the results for the cached chunk whenever it detects that one of its dependencies have changed.

knitr just check if code within .rmd has chenged, not the "CSV" file changes, so we can use this option:

```{r raw_data, cache.extra = file.info("a_very_large_file.csv")}
rawdata <- readr::read_csv("a_very_large_file.csv")
```


**To clean the cache** you can use `r knitr::clean_cache()`.


Global options
------------------------------------
```{r eval = False}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
)
```


we can also use a option for all chanks by default; like when you want just print the output( and not code chanks):

```{r eval = False}
knitr::opts_chunk$set(
  echo = FALSE
)
```

