---
title: "Chapter11 - Colour scales and legends"
output: html_notebook
---

# 11.2 Continuous colour scales


```{r}
erupt <- ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_raster() +
  scale_x_continuous(NULL, expand = c(0, 0)) + 
  scale_y_continuous(NULL, expand = c(0, 0)) + 
  theme(legend.position = "none")
```

## 11.2.1 Particular palettes

```{r}
erupt
erupt + scale_fill_viridis_c()
erupt + scale_fill_viridis_c(option = "magma")
```

```{r}
erupt + scale_fill_distiller()
erupt + scale_fill_distiller(palette = "RdPu")
erupt + scale_fill_distiller(palette = "YlOrBr")
```

```{r}
erupt + scico::scale_fill_scico(palette = "bilbao") # the default
erupt + scico::scale_fill_scico(palette = "vik")
erupt + scico::scale_fill_scico(palette = "lajolla")
```

```{r}
erupt + paletteer::scale_fill_paletteer_c("viridis::plasma")
erupt + paletteer::scale_fill_paletteer_c("scico::tokyo")
erupt + paletteer::scale_fill_paletteer_c("gameofthrones::targaryen")
```

## 11.2.2 Robust recipes

```{r}
erupt
erupt + scale_fill_continuous()
erupt + scale_fill_gradient()
```


```{r}
erupt + scale_fill_gradient(low = "grey", high = "brown")
erupt + scale_fill_gradient2(low = "grey", mid = "white", high = "brown", midpoint = .02)
erupt + scale_fill_gradientn(colours = terrain.colors(7))
```

```{r}
munsell::hue_slice("5P") +  # generate a ggplot with hue_slice()
  annotate(                 # add arrows for annotation 
    geom = "segment", 
    x = c(7, 7), 
    y = c(1, 10), 
    xend = c(7, 7), 
    yend = c(2, 9), 
    arrow = arrow(length = unit(2, "mm"))
  ) 

# construct scale
erupt + scale_fill_gradient(
  low = munsell::mnsl("5P 2/12"), 
  high = munsell::mnsl("5P 7/12")
)
```

```{r}
# munsell example
erupt + scale_fill_gradient2(
  low = munsell::mnsl("5B 7/8"),
  high = munsell::mnsl("5Y 7/8"),
  mid = munsell::mnsl("N 7/0"),
  midpoint = .02
) 

# colorspace examples
erupt + scale_fill_gradientn(colours = colorspace::heat_hcl(7))
erupt + scale_fill_gradientn(colours = colorspace::diverge_hcl(7))
```

## 11.2.3 Missing values
```{r}
df <- data.frame(x = 1, y = 1:5, z = c(1, 3, 2, NA, 5))
base <- ggplot(df, aes(x, y)) + 
  geom_tile(aes(fill = z), size = 5) + 
  labs(x = NULL, y = NULL)

base
base + scale_fill_gradient(na.value = NA)
base + scale_fill_gradient(na.value = "yellow")
```

## 11.2.4 Limits, breaks, and labels

```{r}
toy <- data.frame(
  const = 1, 
  up = 1:4,
  txt = letters[1:4], 
  big = (1:4)*1000,
  log = c(2, 5, 10, 2000)
)
leg <- ggplot(toy, aes(up, up, fill = big)) + 
  geom_tile() + 
  labs(x = NULL, y = NULL) 
leg + scale_fill_continuous(breaks = NULL)
```

# 11.3 Discrete colour scales

```{r}
df <- data.frame(x = c("a", "b", "c", "d"), y = c(3, 4, 1, 2))
bars <- ggplot(df, aes(x, y, fill = x)) + 
  geom_bar(stat = "identity") + 
  labs(x = NULL, y = NULL) +
  theme(legend.position = "none")
```

```{r}
bars
bars + scale_fill_discrete()
bars + scale_fill_hue()
```

## 11.3.1 Brewer scales

```{r}
RColorBrewer::display.brewer.all()
```

```{r}
bars + scale_fill_brewer(palette = "Set1")
bars + scale_fill_brewer(palette = "Set2")
bars + scale_fill_brewer(palette = "Accent")
```

```{r}
# scatter plot
df <- data.frame(x = 1:3 + runif(30), y = runif(30), z = c("a", "b", "c"))
point <- ggplot(df, aes(x, y)) +
  geom_point(aes(colour = z))  + 
  theme(legend.position = "none") +
  labs(x = NULL, y = NULL)

# three palettes
point + scale_colour_brewer(palette = "Set1")
point + scale_colour_brewer(palette = "Set2")  
point + scale_colour_brewer(palette = "Pastel1")
```

```{r}
# bar plot
df <- data.frame(x = 1:3, y = 3:1, z = c("a", "b", "c"))
area <- ggplot(df, aes(x, y)) + 
  geom_bar(aes(fill = z), stat = "identity") + 
  theme(legend.position = "none") +
  labs(x = NULL, y = NULL)

# three palettes
area + scale_fill_brewer(palette = "Set1")
area + scale_fill_brewer(palette = "Set2")
area + scale_fill_brewer(palette = "Pastel1")
```

## 11.3.2 Hue and grey scales

```{r}
bars
bars + scale_fill_hue(c = 40)
bars + scale_fill_hue(h = c(180, 300))
```


```{r}
bars + scale_fill_grey()
bars + scale_fill_grey(start = 0.5, end = 1)
bars + scale_fill_grey(start = 0, end = 0.5)
```

## 11.3.3 Manual scales

```{r}
bars + scale_fill_manual(values = c("sienna1", "sienna4", "hotpink1", "hotpink4"))
bars + scale_fill_manual(values = c("tomato1", "tomato2", "tomato3", "tomato4"))
bars + scale_fill_manual(values = c("grey", "black", "grey", "grey"))
```

```{r}
bars + scale_fill_manual(values = c(
  "d" = "grey",
  "c" = "grey",
  "b" = "black",
  "a" = "grey"
))
```

## 11.3.4 Exercises
### question 1
```{r}
ggplot(mpg, aes(displ, hwy, color = drv)) +
  geom_point() +
  scale_color_discrete(labels = c("4wd", "fwd", "rwd"))
```
# 11.4 Binned colour scales
```{r}
erupt + scale_fill_binned()
```

```{r}
erupt + scale_fill_steps()
```

```{r}
erupt + scale_fill_steps(n.breaks = 8)
```

```{r}
erupt + scale_fill_steps(low = "grey", high = "brown")
```

```{r}
erupt + scale_fill_steps2(low = "grey", mid = "white", high = "brown", midpoint = .02)
```

```{r}
erupt + scale_fill_stepsn(n.breaks = 12, colours = terrain.colors(12))
```

```{r}
erupt + scale_fill_fermenter(n.breaks = 9)
```

```{r}
erupt + scale_fill_fermenter(n.breaks = 9, palette = "Oranges")
```

```{r}
erupt + scale_fill_fermenter(n.breaks = 9, palette = "PuOr")
```

# 11.6 Legends

## 11.6.1 Positioning

```{r}
base <- ggplot(toy, aes(up, up)) + 
  geom_point(aes(colour = txt), size = 3) + 
  xlab(NULL) + 
  ylab(NULL)

base + theme(legend.position = "left")
base + theme(legend.position = "right") # the default 
base + theme(legend.position = "bottom")
base + theme(legend.position = "none")
```

```{r}
base <- ggplot(toy, aes(up, up)) + 
  geom_point(aes(colour = txt), size = 3)

base + theme(legend.position = c(0, 1), legend.justification = c(0, 1))
base + theme(legend.position = c(0.5, 0.5), legend.justification = c(0.5, 0.5))
base + theme(legend.position = c(1, 0), legend.justification = c(1, 0))
```
There’s also a margin around the legends, which you can suppress with `legend.margin = unit(0, "mm")`.


# 11.7 Legend key glyphs

```{r}
base <- ggplot(economics, aes(date, psavert, color = "savings"))

base + geom_line()
base + geom_line(key_glyph = "timeseries")
```

## 11.7.1 guide_legend()

```{r}
base <- ggplot(mpg, aes(drv, fill = factor(cyl))) + geom_bar() 

base
base + guides(fill = guide_legend(ncol = 2))
base + guides(fill = guide_legend(ncol = 2, byrow = TRUE))
```

```{r}
base
base + guides(fill = guide_legend(reverse = TRUE))
```

```{r}
base
base + guides(fill = guide_legend(reverse = TRUE))
```

## 11.7.2 guide_bins()

```{r}
base <- ggplot(mpg, aes(displ, manufacturer, size = hwy)) +
  geom_point(alpha = .2) +
  scale_size_binned()
```

```{r}
base + guides(size = guide_bins(direction = "vertical"))
base + guides(size = guide_bins(direction = "horizontal"))
```

```{r}
base + guides(size = guide_bins(show.limits = TRUE))
base + guides(
  size = guide_bins(
    axis.colour = "red",
    axis.arrow = arrow(
      length = unit(.1, "inches"), 
      ends = "first", 
      type = "closed"
    )
  )
)
```

## 11.7.3 guide_colourbar() / guide_colorbar()

```{r}
base <- ggplot(mpg, aes(cyl, displ, colour = hwy)) +
  geom_point(size = 2)

base
base + guides(colour = guide_colourbar(reverse = TRUE))
base + guides(colour = guide_colourbar(barheight = unit(2, "cm")))
```

## 11.7.4 guide_coloursteps() / guide_colorsteps()

```{r}
base <- ggplot(mpg, aes(displ, hwy, colour = cyl)) + 
  geom_point() + 
  scale_color_binned()

base + guides(colour = guide_coloursteps(show.limits = TRUE))
base + guides(colour = guide_coloursteps(show.limits = FALSE))
```

## 11.7.5 Exercises

### question 1

```{r}
ggplot(toy, aes(up, up)) + 
  geom_point(aes(colour = txt), size = 3) + 
  xlab(NULL) + 
  ylab(NULL) +
  theme(legend.position = "left")
```

### question 2

```{r}
ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(color = drv, shape = drv)) +
  labs(color = "Drive train", shape = "Drive train")
```


### question 3
```{r}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point(key_glyph = draw_key_path) +
  geom_smooth(method = "lm", se = F) +
  theme(legend.position = "bottom")
```

