---
title: "Chapter5 - Case study: ER injuries"
output: html_notebook
---


```{r setup}
knitr::opts_chunk$set(eval = FALSE)
```


```{r}
library(shiny)
library(vroom)
library(tidyverse)
library(neiss)
```


creating `.tsv` files:
```{r}
top_prod <- injuries %>% 
  filter(trmt_date >= as.Date("2017-01-01"), trmt_date < as.Date("2018-01-01")) %>% 
  count(prod1, sort = TRUE) %>% 
  filter(n > 5 * 365)

injuries %>%
  filter(trmt_date >= as.Date("2017-01-01"), trmt_date < as.Date("2018-01-01")) %>%
  semi_join(top_prod, by = "prod1") %>%
  mutate(age = floor(age), sex = tolower(sex), race = tolower(race)) %>%
  filter(sex != "unknown") %>%
  select(trmt_date, age, sex, race, body_part, diag, location, prod_code = prod1, weight, narrative) %>%
  vroom::vroom_write("neiss/injuries.tsv.gz")

products %>%
  semi_join(top_prod, by = c("code" = "prod1")) %>%
  rename(prod_code = code) %>%
  vroom::vroom_write("neiss/products.tsv")

population %>%
  filter(year == 2017) %>%
  select(-year) %>%
  rename(population = n) %>%
  vroom::vroom_write("neiss/population.tsv")
```

# 5.2 The data


```{r}
injuries <- vroom::vroom("neiss/injuries.tsv.gz")
injuries
```

```{r}
products <- vroom::vroom("neiss/products.tsv")
products
population <- vroom::vroom("neiss/population.tsv")
population
```

# 5.3 Exploration

```{r}
selected <- injuries %>% filter(prod_code == 649)
nrow(selected)
selected %>% count(location, wt = weight, sort = TRUE)
selected %>% count(body_part, wt = weight, sort = TRUE)
selected %>% count(diag, wt = weight, sort = TRUE)
```



```{r}
summary <- selected %>% 
  count(age, sex, wt = weight)
summary

summary %>% 
  ggplot(aes(age, n, colour = sex)) + 
  geom_line() + 
  labs(y = "Estimated number of injuries")
```

```{r}
summary <- selected %>% 
  count(age, sex, wt = weight) %>% 
  left_join(population, by = c("age", "sex")) %>% 
  mutate(rate = n / population * 1e4)

summary
```
```{r}
summary %>% 
  ggplot(aes(age, rate, colour = sex)) + 
  geom_line(na.rm = TRUE) + 
  labs(y = "Injuries per 10,000 people")
```

```{r}
selected %>% 
  sample_n(10) %>% 
  pull(narrative)
```

# 5.4 Prototype

> When designing a first prototype, the challenge is in making it “as simple as possible”. There’s a tension between getting the basics working quickly and planning for the future of the app. To help get the balance right, I often do a few pencil-and-paper sketches to rapidly explore the UI and reactive graph before committing to code.


```{r}
ui <- fluidPage(
  fluidRow(
    column(6,
      selectInput("code", "Product", setNames(products$prod_code, products$title))
    )
  ),
  fluidRow(
    column(4, tableOutput("diag")),
    column(4, tableOutput("body_part")),
    column(4, tableOutput("location"))
  ),
  fluidRow(
    column(12, plotOutput("age_sex"))
  )
)

server <- function(input, output, session) {
  selected <- reactive(injuries %>% filter(prod_code == input$code))

  output$diag <- renderTable(
    selected() %>% count(diag, wt = weight, sort = TRUE)
  )
  output$body_part <- renderTable(
    selected() %>% count(body_part, wt = weight, sort = TRUE)
  )
  output$location <- renderTable(
    selected() %>% count(location, wt = weight, sort = TRUE)
  )

  summary <- reactive({
    selected() %>%
      count(age, sex, wt = weight) %>%
      left_join(population, by = c("age", "sex")) %>%
      mutate(rate = n / population * 1e4)
  })

  output$age_sex <- renderPlot({
    summary() %>%
      ggplot(aes(age, n, colour = sex)) +
      geom_line() +
      labs(y = "Estimated number of injuries")
  }, res = 96)
}
shinyApp(ui, server)
```


# 5.5 Polish tables

```{r}
injuries %>%
  mutate(diag = fct_lump(fct_infreq(diag), n = 5)) %>%
  group_by(diag) %>%
  summarise(n = as.integer(sum(weight)))
```


```{r}
count_top <- function(df, var, n = 5) {
  df %>%
    mutate({{ var }} := fct_lump(fct_infreq({{ var }}), n = n)) %>%
    group_by({{ var }}) %>%
    summarise(n = as.integer(sum(weight)))
}
```


```{r}
server <- function(input, output, session){
  selected <- reactive(injuries %>% filter(prod_code == input$code))
  output$diag <- renderTable(count_top(selected(), diag), width = "100%")
  output$body_part <- renderTable(count_top(selected(), body_part), width = "100%")
  output$location <- renderTable(count_top(selected(), location), width = "100%")
  summary <- reactive({
    selected() %>%
      count(age, sex, wt = weight) %>%
      left_join(population, by = c("age", "sex")) %>%
      mutate(rate = n / population * 1e4)
  })

  output$age_sex <- renderPlot({
    summary() %>%
      ggplot(aes(age, n, colour = sex)) +
      geom_line() +
      labs(y = "Estimated number of injuries")
  }, res = 96)
}

shinyApp(ui, server)
```

```{r}
ui <- fluidPage(
  fluidRow(
    column(8,
      selectInput("code", "Product",
        choices = setNames(products$prod_code, products$title),
        width = "100%"
      )
    ),
    column(2, selectInput("y", "Y axis", c("rate", "count")))
  ),
  fluidRow(
    column(4, tableOutput("diag")),
    column(4, tableOutput("body_part")),
    column(4, tableOutput("location"))
  ),
  fluidRow(
    column(12, plotOutput("age_sex"))
  )
)

server <- function(input, output, session){
  selected <- reactive(injuries %>% filter(prod_code == input$code))
  output$diag <- renderTable(count_top(selected(), diag), width = "100%")
  output$body_part <- renderTable(count_top(selected(), body_part), width = "100%")
  output$location <- renderTable(count_top(selected(), location), width = "100%")
  summary <- reactive({
    selected() %>%
      count(age, sex, wt = weight) %>%
      left_join(population, by = c("age", "sex")) %>%
      mutate(rate = n / population * 1e4)
  })

   output$age_sex <- renderPlot({
    if (input$y == "count") {
      summary() %>%
        ggplot(aes(age, n, colour = sex)) +
        geom_line() +
        labs(y = "Estimated number of injuries")
    } else {
      summary() %>%
        ggplot(aes(age, rate, colour = sex)) +
        geom_line(na.rm = TRUE) +
        labs(y = "Injuries per 10,000 people")
    }
  }, res = 96)

}
shinyApp(ui, server)
``` 

# 5.7 Narrative

```{r}
ui <- fluidPage(
  fluidRow(
    column(8,
      selectInput("code", "Product",
        choices = setNames(products$prod_code, products$title),
        width = "100%"
      )
    ),
    column(2, selectInput("y", "Y axis", c("rate", "count")))
  ),
  fluidRow(
    column(4, tableOutput("diag")),
    column(4, tableOutput("body_part")),
    column(4, tableOutput("location"))
  ),
  fluidRow(
    column(12, plotOutput("age_sex"))
  ),
    fluidRow(
    column(2, actionButton("story", "Tell me a story")),
    column(10, textOutput("narrative"))
  )
)

server <- function(input, output, session){
  selected <- reactive(injuries %>% filter(prod_code == input$code))
  output$diag <- renderTable(count_top(selected(), diag), width = "100%")
  output$body_part <- renderTable(count_top(selected(), body_part), width = "100%")
  output$location <- renderTable(count_top(selected(), location), width = "100%")
  summary <- reactive({
    selected() %>%
      count(age, sex, wt = weight) %>%
      left_join(population, by = c("age", "sex")) %>%
      mutate(rate = n / population * 1e4)
  })

   output$age_sex <- renderPlot({
    if (input$y == "count") {
      summary() %>%
        ggplot(aes(age, n, colour = sex)) +
        geom_line() +
        labs(y = "Estimated number of injuries")
    } else {
      summary() %>%
        ggplot(aes(age, rate, colour = sex)) +
        geom_line(na.rm = TRUE) +
        labs(y = "Injuries per 10,000 people")
    }
  }, res = 96)
   
   
 output$narrative <- renderText({
    input$story
    selected() %>% pull(narrative) %>% sample(1)
  })
}
shinyApp(ui, server)
```

# 5.8 Exercises

### question 2
```{r}
diag <- injuries %>%
  mutate(diag = fct_lump(fct_infreq(diag), n = 5)) %>%
  pull(diag)

levels(diag)


diag <- injuries %>%
  mutate(diag = fct_infreq(fct_lump(diag, n = 5))) %>%
  pull(diag)

levels(diag)
```


### question 3


```{r}

ui <- fluidPage(
  fluidRow(
    column(8,
      selectInput("code", "Product",
        choices = setNames(products$prod_code, products$title),
        width = "100%"
      )
    ),
    column(2, selectInput("y", "Y axis", c("rate", "count"))),
    column(2, numericInput("nrow", "# rows", value = 5, min = 1 , max = 10))
  ),
  fluidRow(
    column(4, tableOutput("diag")),
    column(4, tableOutput("body_part")),
    column(4, tableOutput("location"))
  ),
  fluidRow(
    column(12, plotOutput("age_sex"))
  ),
    fluidRow(
    column(2, actionButton("story", "Tell me a story")),
    column(10, textOutput("narrative"))
  )
)

server <- function(input, output, session){
  selected <- reactive(injuries %>% filter(prod_code == input$code))
  
    max_no_rows <- reactive(
    max(length(unique(selected()$diag)),
        length(unique(selected()$body_part)),
        length(unique(selected()$location)))
  )
  observeEvent(input$code, {
    updateNumericInput(session, "nrow", max = max_no_rows())
  })
  table_rows <- reactive(input$nrow)
  output$diag <- renderTable(count_top(selected(), diag, n = table_rows()), width = "100%")
  output$body_part <- renderTable(count_top(selected(), body_part, n = table_rows()), width = "100%")
  output$location <- renderTable(count_top(selected(), location, n = table_rows()), width = "100%")
  summary <- reactive({
    selected() %>%
      count(age, sex, wt = weight) %>%
      left_join(population, by = c("age", "sex")) %>%
      mutate(rate = n / population * 1e4)
  })
   
   output$age_sex <- renderPlot({
    if (input$y == "count") {
      summary() %>%
        ggplot(aes(age, n, colour = sex)) +
        geom_line() +
        labs(y = "Estimated number of injuries")
    } else {
      summary() %>%
        ggplot(aes(age, rate, colour = sex)) +
        geom_line(na.rm = TRUE) +
        labs(y = "Injuries per 10,000 people")
    }
  }, res = 96)
   
   
 output$narrative <- renderText({
    input$story
    selected() %>% pull(narrative) %>% sample(1)
  })
}
shinyApp(ui, server)
```


### question 4

```{r}
ui <- fluidPage(
  fluidRow(
    column(8,
      selectInput("code", "Product",
        choices = setNames(products$prod_code, products$title),
        width = "100%"
      )
    ),
    column(2, selectInput("y", "Y axis", c("rate", "count"))),
    column(2, numericInput("nrow", "# rows", value = 5, min = 1 , max = 10))
  ),
  fluidRow(
    column(4, tableOutput("diag")),
    column(4, tableOutput("body_part")),
    column(4, tableOutput("location"))
  ),
  fluidRow(
    column(12, plotOutput("age_sex"))
  ),
    fluidRow(
    column(2, actionButton("prev_story", "Previous story")),
    column(8, textOutput("narrative")),
    column(2, actionButton("next_story", "Next story")),
  )
)


index_val <- 1
server <- function(input, output, session){
  selected <- reactive(injuries %>% filter(prod_code == input$code))
  
    max_no_rows <- reactive(
    max(length(unique(selected()$diag)),
        length(unique(selected()$body_part)),
        length(unique(selected()$location)))
  )
  observeEvent(input$code, {
    updateNumericInput(session, "nrow", max = max_no_rows())
  })
  table_rows <- reactive(input$nrow)
  output$diag <- renderTable(count_top(selected(), diag, n = table_rows()), width = "100%")
  output$body_part <- renderTable(count_top(selected(), body_part, n = table_rows()), width = "100%")
  output$location <- renderTable(count_top(selected(), location, n = table_rows()), width = "100%")
  summary <- reactive({
    selected() %>%
      count(age, sex, wt = weight) %>%
      left_join(population, by = c("age", "sex")) %>%
      mutate(rate = n / population * 1e4)
  })
   
   output$age_sex <- renderPlot({
    if (input$y == "count") {
      summary() %>%
        ggplot(aes(age, n, colour = sex)) +
        geom_line() +
        labs(y = "Estimated number of injuries")
    } else {
      summary() %>%
        ggplot(aes(age, rate, colour = sex)) +
        geom_line(na.rm = TRUE) +
        labs(y = "Injuries per 10,000 people")
    }
  }, res = 96)
   
   
   max_len <- reactive(length(selected()$narrative))
   
   
   next_index <- eventReactive(input$next_story,{
     index_val + 1
   })
   
   prev_index <- eventReactive(input$prev_story,{
     index_val - 1
   })
   observeEvent(input$next_story, {
     print(next_index())
     index_val <<- next_index()
   })
   
   observeEvent(input$prev_story, {
     print(prev_index())
     index_val <<- prev_index()
   })
   
   output$narrative <- renderText({
      selected()$narrative[index_val]
   })
}
shinyApp(ui, server)
```

