---
title: "Chapter19 - Shiny modules"
output: html_notebook
---

```{r setup}
knitr::opts_chunk$set(eval = FALSE)
```


```{r}
library(shiny)
```


# 19.2 Module basics

```{r}
ui <- fluidPage(
  selectInput("var", "Variable", names(mtcars)),
  numericInput("bins", "bins", 10, min = 1),
  plotOutput("hist")
)
server <- function(input, output, session) {
  data <- reactive(mtcars[[input$var]])
  output$hist <- renderPlot({
    hist(data(), breaks = input$bins, main = input$var)
  }, res = 96)
}
```


## 19.2.1 Module UI


Here I’ve returned the UI components in a `tagList()`, but you could also return them in an HTML container like `column()` or a `fluidRow()`. Returning a list is more flexible because it allows the caller of the module to choose the container. If you always place the module in the same container, you should return that instead, saving a little code in the UI.


```{r}
histogramUI <- function(id) {
  tagList(
    selectInput(NS(id, "var"), "Variable", names(mtcars)),
    numericInput(NS(id, "bins"), "bins", 10, min = 1),
    plotOutput(NS(id, "hist"))
  )
}
```




## 19.2.2 Module server

```{r}
histogramServer <- function(id) {
  moduleServer(id, function(input, output, session) {
    data <- reactive(mtcars[[input$var]])
    output$hist <- renderPlot({
      hist(data(), breaks = input$bins, main = input$var)
    }, res = 96)
  })
}
```


## 19.2.3 Updated app

```{r}
histogramApp <- function() {
  ui <- fluidPage(
    histogramUI("hist1")
  )
  server <- function(input, output, session) {
    histogramServer("hist1")
  }
  shinyApp(ui, server)  
}
histogramApp()
```

## 19.2.4 Namespacing

```{r}
ui <- fluidPage(
  histogramUI("hist1"),
  textOutput("out")
)
server <- function(input, output, session) {
  histogramServer("hist1")
  output$out <- renderText(paste0("Bins: ", input$bins))
}
shinyApp(ui, server)
```


# 19.3 Inputs and outputs

## 19.3.1 Getting started: UI input + server output

```{r}
datasetInput <- function(id, filter = NULL) {
  names <- ls("package:datasets")
  if (!is.null(filter)) {
    data <- lapply(names, get, "package:datasets")
    names <- names[vapply(data, filter, logical(1))]
  }
  
  selectInput(NS(id, "dataset"), "Pick a dataset", choices = names)
}
```

```{r}
datasetServer <- function(id) {
  moduleServer(id, function(input, output, session) {
    reactive(get(input$dataset, "package:datasets"))
  })
}
```

```{r}
datasetApp <- function(filter = NULL) {
  ui <- fluidPage(
    datasetInput("dataset", filter = filter),
    tableOutput("data")
  )
  server <- function(input, output, session) {
    data <- datasetServer("dataset")
    output$data <- renderTable(head(data()))
  }
  shinyApp(ui, server)
}
datasetApp(is.data.frame)
```

## 19.3.2 Case study: selecting a numeric variable

```{r}
selectVarInput <- function(id) {
  selectInput(NS(id, "var"), "Variable", choices = NULL) 
}


find_vars <- function(data, filter) {
  names(data)[vapply(data, filter, logical(1))]
}

selectVarServer <- function(id, data, filter = is.numeric) {
  moduleServer(id, function(input, output, session) {
    observeEvent(data(), {
      updateSelectInput(session, "var", choices = find_vars(data(), filter))
    })
    
    reactive(data()[[input$var]])
  })
}
```


```{r}
selectVarApp <- function(filter = is.numeric) {
  ui <- fluidPage(
    datasetInput("data", is.data.frame),
    selectVarInput("var"),
    verbatimTextOutput("out")
  )
  server <- function(input, output, session) {
    data <- datasetServer("data")
    var <- selectVarServer("var", data, filter = filter)
    output$out <- renderPrint(var())
  }
  
  shinyApp(ui, server)
}
selectVarApp()
```

## 19.3.3 Server inputs
```{r}
stopifnot(is.reactive(data))
stopifnot(!is.reactive(filter))
```

```{r}
find_vars <- function(data, filter) {
  stopifnot(is.data.frame(data))
  stopifnot(is.function(filter))
  names(data)[vapply(data, filter, logical(1))]
}
```


## 19.3.4 Modules inside of modules

```{r}
selectDataVarUI <- function(id) {
  tagList(
    datasetInput(NS(id, "data"), filter = is.data.frame),
    selectVarInput(NS(id, "var"))
  )
}
selectDataVarServer <- function(id, filter = is.numeric) {
  moduleServer(id, function(input, output, session) {
    data <- datasetServer("data")
    var <- selectVarServer("var", data, filter = filter)
    var
  })
}

selectDataVarApp <- function(filter = is.numeric) {
  ui <- fluidPage(
    sidebarLayout(
      sidebarPanel(selectDataVarUI("var")),
      mainPanel(verbatimTextOutput("out"))
    )
  )
  server <- function(input, output, session) {
    var <- selectDataVarServer("var", filter)
    output$out <- renderPrint(var(), width = 40)
  }
  shinyApp(ui, server)
}
```

## 19.3.5 Case study: histogram

```{r}
histogramOutput <- function(id) {
  tagList(
    numericInput(NS(id, "bins"), "bins", 10, min = 1, step = 1),
    plotOutput(NS(id, "hist"))
  )
}

histogramServer <- function(id, x, title = reactive("Histogram")) {
  stopifnot(is.reactive(x))
  stopifnot(is.reactive(title))
  
  moduleServer(id, function(input, output, session) {
    output$hist <- renderPlot({
      req(is.numeric(x()))
      main <- paste0(title(), " [", input$bins, "]")
      hist(x(), breaks = input$bins, main = main)
    }, res = 96)
  })
}
```

```{r}
histogramApp <- function() {
  ui <- fluidPage(
    sidebarLayout(
      sidebarPanel(
        datasetInput("data", is.data.frame),
        selectVarInput("var"),
      ),
      mainPanel(
        histogramOutput("hist")    
      )
    )
  )
  
  server <- function(input, output, session) {
    data <- datasetServer("data")
    x <- selectVarServer("var", data)
    histogramServer("hist", x)
  }
  shinyApp(ui, server)
} 
histogramApp()
```

```{r}
histogramOutputBins <- function(id) {
  numericInput(NS(id, "bins"), "bins", 10, min = 1, step = 1)
}
histogramOutputPlot <- function(id) {
  plotOutput(NS(id, "hist"))
}

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      datasetInput("data", is.data.frame),
      selectVarInput("var"),
      histogramOutputBins("hist")
    ),
    mainPanel(
      histogramOutputPlot("hist")
    )
  )
)
```

## 19.3.6 Multiple outputs

```{r}
selectVarServer <- function(id, data, filter = is.numeric) {
  stopifnot(is.reactive(data))
  stopifnot(!is.reactive(filter))
  
  moduleServer(id, function(input, output, session) {
    observeEvent(data(), {
      updateSelectInput(session, "var", choices = find_vars(data(), filter))
    })
    
    list(
      name = reactive(input$var),
      value = reactive(data()[[input$var]])
    )
  })
}
```

```{r}
histogramApp <- function() {
  ui <- fluidPage(
    sidebarLayout(
      sidebarPanel(
        datasetInput("data", is.data.frame),
        selectVarInput("var"),
      ),
      mainPanel(
        histogramOutput("hist")    
      )
    )
  )

  server <- function(input, output, session) {
    data <- datasetServer("data")
    x <- selectVarServer("var", data)
    histogramServer("hist", x$value, x$name)
  }
  shinyApp(ui, server)
} 
```

```{r}
 server <- function(input, output, session) {
    data <- datasetServer("data")
    c(value, name) %<-% selectVarServer("var", data)
    histogramServer("hist", value, name)
  }
```


# 19.4 Case studies

## 19.4.1 Limited selection + other

```{r}
radioExtraUI <- function(id, label, choices, selected = NULL, placeholder = "Other") {
  other <- textInput(NS(id, "other"), label = NULL, placeholder = placeholder)
  
  names <- if (is.null(names(choices))) choices else names(choices)
  values <- unname(choices)
  
  radioButtons(NS(id, "primary"), 
    label = label,
    choiceValues = c(names, "other"),
    choiceNames = c(as.list(values), list(other)),
    selected = selected
  )
}
```

```{r}
radioExtraServer <- function(id) {
  moduleServer(id, function(input, output, session) {
    observeEvent(input$other, ignoreInit = TRUE, {
      updateRadioButtons(session, "primary", selected = "other")
    })
    
    reactive({
      if (input$primary == "other") {
        input$other
      } else {
        input$primary
      }
    })
  })
}
```


```{r}
radioExtraApp <- function(...) {
  ui <- fluidPage(
    radioExtraUI("extra", NULL, ...),
    textOutput("value")
  )
  server <- function(input, output, server) {
    extra <- radioExtraServer("extra")
    output$value <- renderText(paste0("Selected: ", extra()))
  }
  
  shinyApp(ui, server)
}
radioExtraApp(c("a", "b", "c"))
```


```{r}
genderUI <- function(id, label = "Gender") {
  radioExtraUI(id, 
    label = label,
    choices = c(
      male = "Male",
      female = "Female",
      na = "Prefer not to say"
    ), 
    placeholder = "Self-described", 
    selected = "na"
  )
}
```


## 19.4.2 Wizard


```{r}
nextPage <- function(id, i) {
  actionButton(NS(id, paste0("go_", i, "_", i + 1)), "next")
}
prevPage <- function(id, i) {
  actionButton(NS(id, paste0("go_", i, "_", i - 1)), "prev")
}
```

```{r}
wrapPage <- function(title, page, button_left = NULL, button_right = NULL) {
  tabPanel(
    title = title, 
    fluidRow(
      column(12, page)
    ), 
    fluidRow(
      column(6, button_left),
      column(6, button_right)
    )
  )
}
```


```{r}
wizardUI <- function(id, pages, doneButton = NULL) {
  stopifnot(is.list(pages))
  n <- length(pages)
  
  wrapped <- vector("list", n)
  for (i in seq_along(pages)) {
    # First page only has next; last page only prev + done
    lhs <- if (i > 1) prevPage(id, i)
    rhs <- if (i < n) nextPage(id, i) else doneButton
    wrapped[[i]] <- wrapPage(paste0("page_", i), pages[[i]], lhs, rhs)
  }
  
  # Create tabsetPanel
  # https://github.com/rstudio/shiny/issues/2927
  wrapped$id <- NS(id, "wizard")
  wrapped$type <- "hidden"
  do.call("tabsetPanel", wrapped)
}
```

```{r}
wizardServer <- function(id, n) {
  moduleServer(id, function(input, output, session) {
    changePage <- function(from, to) {
      observeEvent(input[[paste0("go_", from, "_", to)]], {
        updateTabsetPanel(session, "wizard", selected = paste0("page_", to))
      })  
    }
    ids <- seq_len(n)
    lapply(ids[-1], function(i) changePage(i, i - 1))
    lapply(ids[-n], function(i) changePage(i, i + 1))
  })
}
```

```{r}
wizardApp <- function(...) {
  pages <- list(...)
  
  ui <- fluidPage(
    wizardUI("whiz", pages)
  )
  server <- function(input, output, session) {
    wizardServer("whiz", length(pages))
  }
  shinyApp(ui, server)
}
wizardApp(
  "Page 1",
  "Page 2",
  "Page 3"
)
```


```{r}
page1 <- tagList(
  textInput("name", "What's your name?")
)
page2 <- tagList(
  numericInput("age", "How old are you?", 20)
)
page3 <- tagList(
  "Is this data correct?",
  verbatimTextOutput("info")
)

ui <- fluidPage(
  wizardUI(
    id = "demographics", 
    pages = list(page1, page2, page3), 
    doneButton = actionButton("done", "Submit")
  )
)
server <- function(input, output, session) {
  wizardServer("demographics", 3)

  observeEvent(input$done, showModal(
    modalDialog("Thank you!", footer = NULL)
  ))
  
  output$info <- renderText(paste0(
    "Age: ", input$age, "\n",
    "Name: ", input$name, "\n"
  ))
}
shinyApp(ui, server)
```


## 19.4.3 Dynamic UI


```{r}
filterUi <- function(id) {
  uiOutput(NS(id, "controls"))
}
```

```{r}
library(purrr)

make_ui <- function(x, id, var) {
  if (is.numeric(x)) {
    rng <- range(x, na.rm = TRUE)
    sliderInput(id, var, min = rng[1], max = rng[2], value = rng)
  } else if (is.factor(x)) {
    levs <- levels(x)
    selectInput(id, var, choices = levs, selected = levs, multiple = TRUE)
  } else {
    # Not supported
    NULL
  }
}
filter_var <- function(x, val) {
  if (is.numeric(x)) {
    !is.na(x) & x >= val[1] & x <= val[2]
  } else if (is.factor(x)) {
    x %in% val
  } else {
    # No control, so don't filter
    TRUE
  }
}
```


```{r}
filterServer <- function(id, df) {
  stopifnot(is.reactive(df))

  moduleServer(id, function(input, output, session) {
    vars <- reactive(names(df()))
    
    output$controls <- renderUI({
      map(vars(), function(var) make_ui(df()[[var]], NS(id, var), var))
    })
    
    reactive({
      each_var <- map(vars(), function(var) filter_var(df()[[var]], input[[var]]))
      reduce(each_var, `&`)
    })
  })
}
```


```{r}
filterApp <- function() {
  ui <- fluidPage(
    sidebarLayout(
      sidebarPanel(
        datasetInput("data", is.data.frame),
        textOutput("n"),
        filterUi("filter"),
      ),
      mainPanel(
        tableOutput("table")    
      )
    )
  )
  server <- function(input, output, session) {
    df <- datasetServer("data")
    filter <- filterServer("filter", df)
    
    output$table <- renderTable(df()[filter(), , drop = FALSE])
    output$n <- renderText(paste0(sum(filter()), " rows"))
  }
  shinyApp(ui, server)
}
filterApp()
```

# 19.5 Single object modules

```{r}
histogramUI <- function(id, df) {
  tagList(
    selectInput(NS(id, "var"), "Variable", names(df)),
    numericInput(NS(id, "bins"), "bins", 10, min = 1),
    plotOutput(NS(id, "hist"))
  )
}

histogramServer <- function(id, df) {
  moduleServer(id, function(input, output, session) {
    data <- reactive(df[[input$var]])
    output$hist <- renderPlot({
      hist(data(), breaks = input$bins, main = input$var)
    }, res = 96)
  })
}
```

```{r}
ui <- fluidPage(
  tabsetPanel(
    tabPanel("mtcars", histogramUI("mtcars", mtcars)),
    tabPanel("iris", histogramUI("iris", iris))
  )
)
server <- function(input, output, session) {
  histogramServer("mtcars", mtcars)
  histogramServer("iris", iris)
}
```


```{r}
histogramApp <- function(id, df) {
  list(
    ui = histogramUI(id, df), 
    server = histogramServer(id, df)
  )
}
```

```{r}
hist1 <- histogramApp("mtcars", mtcars)
hist2 <- histogramApp("iris", iris)

ui <- fluidPage(
  tabsetPanel(
    tabPanel("mtcars", hist1$ui()),
    tabPanel("iris", hist2$ui())
  )
)
server <- function(input, output, session) {
  hist1$server()
  hist2$server()
}
```

# 19.6 Exercises

### question 2

```{r}
summaryOutput <- function(id) {
  tags$ul(
    tags$li("Min: ", textOutput(NS(id, "min"), inline = TRUE)),
    tags$li("Max: ", textOutput(NS(id, "max"), inline = TRUE)),
    tags$li("Missing: ", textOutput(NS(id, "n_na"), inline = TRUE))
  )
}

summaryServer <- function(id, var) {
  moduleServer(id, function(input, output, session) {
    rng <- reactive({
      req(var())
      range(var(), na.rm = TRUE)
    })

    output$min <- renderText(rng()[[1]])
    output$max <- renderText(rng()[[2]])
    output$n_na <- renderText(sum(is.na(var())))
  })
}
```

```{r}
selectVarApp <- function(filter = is.numeric) {
    ui <- fluidPage(
        datasetInput("data", is.data.frame),
        selectVarInput("var"),
        summaryOutput("summary")
    )
    server <- function(input, output, session) {
        data <- datasetServer("data")
        var <- selectVarServer("var", data, filter = filter)
        summaryServer("summary", var)
    }

    shinyApp(ui, server)
}

selectVarApp()
```
